Predicting the Likelihood of Climate-Friendly Actions: Insights from Age and Educational Attainment\thanks{Code and data are available at: \url{https://github.com/LexiKnight/toronto_climate/tree/main}.}

### 2024-12-11 paper/climate_change.tex
save models into correct model folder

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Barriers to Green Living: Insights from 2018 and 2021 Toronto Climate Perception Surveys},
  pdfauthor={Lexi Knight},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Barriers to Green Living: Insights from 2018 and 2021 Toronto
Climate Perception Surveys\thanks{Code and data are available at:
\url{https://github.com/LexiKnight/toronto_climate/tree/main}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Uncovering What Stops Sustainable Behaviors and How
Communities Can Overcome Them}
\author{Lexi Knight}
\date{December 3, 2024}

@@

The \texttt{arrow} package was employed to handle large datasets
efficiently, ensuring smooth data processing, while \texttt{forcats} was
used for reordering factor levels to facilitate more insightful analyses
and visualizations. In terms of data tidying, the \texttt{tidyr} and
\texttt{stringr} packages allowed us to reshape and clean the data by
managing character strings and handling missing values. For the final
documentation and report generation, the \texttt{knitr} package was
instrumental in integrating R code and outputs into a cohesive and
reproducible report format. This suite of R packages collectively
enabled us to conduct thorough data management, manipulation, and
analysis, ultimately contributing to the successful completion of the
project.

@@

(\textbf{table-one?}) illustrates the different age ditribution for 2018
and 2021 surveys. The major difference is that there was greater
representation of younger inviduals ranging from 18 to 23 in 2021
compared to 2018. The remaining age groups were not very different.

@@

This table highlights the age distribution of survey respondents in 2018
and 2021.

@@

Figure~\ref{fig-two} illustrates the distribution of survey respondents
by their highest level of education attained. The most frequently
reported education levels are ``Completed undergraduate degree'' and
``Post graduate/professional school,'' with both categories having a
noticeably higher frequency compared to others, indicating a
well-educated respondent population. In contrast, fewer respondents
reported ``Some community college, vocational, trade school'' and
``Completed community college, vocational, trade school,'' suggesting
these levels are less common among participants. Additionally, a small
portion of respondents preferred not to disclose their education level.
This distribution highlights a skew toward higher educational attainment
within the sample, which may influence responses related to
socio-economic factors and other survey topics.

@@

\caption{\label{fig-two}The histogram shows the distribution of
education levels of 2018 survey respondents.}

@@

(\textbf{table-two?}) illustrates respondant education level in 2018 and
2021 respectively. Please note, the categories of education level are
different for 2018 and 2021 as these were different surveys and
categories were updated in 2021. Despite this, we are still able to make
comparisons. Individuals with high school diplomas or less account for 9
for 2018 and 38 for 2021 (addition of less than highs school and high
school or equivalent categories). This could indicate that the younger
population was included more in the 2021 suvey or that less education
individuals were surveyed. Additionally, its not super clear howver
those with a post graduate or professional degree accounted for 29
percent in 2018 and 35 percent in 2021. This illustrates that more
educated inviduals responded to the 2021 survey.

@@

This table highlights the education distribution of survey respondents
in 2018 and 2021.''

@@

Figure~\ref{fig-three} depicts the distribution of self-reported climate
change knowledge among survey respondents in 2018. The majority of
participants identified as ``Very informed,'' making this the most
prevalent category by a significant margin. A moderate number of
respondents reported being ``Not very informed,'' while a smaller yet
noticeable group described themselves as ``Extremely informed.'' In
contrast, the categories ``Not at all informed'' and ``Not very
informed'' have the lowest frequencies, with ``Not at all informed''
being particularly rare. This distribution suggests that most
respondents possess a relatively high level of awareness about climate
change, with very few indicating a lack of knowledge on the topic.

@@

\caption{\label{fig-three}The histogram shows the distribution of
self-reported climate change knowledge among 2018 survey respondents.}

@@

(\textbf{table-three?}) illustrates the self-reported level of extent
inviduals feel informed about climate change and climate action for 2018
and 2021. In 2018, the large marjority of individuals say they are very
informed with 60 percent and the follwoign category is not very informed
with 27 percent. In 2021, The majority of inividuals also say they are
very informed but to a lesser degree with 48 percent and this is closely
followed by only a little informed with 37 percent. There is a trend
that invidiusls in 2018 reported greater levels of knowldege about
climate change whereas 2021 illstrates a more even split between very
informed and only little informed, portraying that fewer individuals
have adequate knowledge about climate change.

@@

This table highlights the extent feel informed distribution of survey
respondents in 2018 and 2021.

@@

The analysis of the best method for delivering climate change
information portrayed in Figure~\ref{fig-four} revealed that the
majority of individuals in the 2018 dataset preferred receiving
information firstly from advertising campaigns with about 55 percent,
newsletters with about 52 percent, followed by the toronto website wiht
about 46 percent. On the other side of the spectrum, few said they were
uninterested in receiving climate action information with about 5
percent, about 12 percent indicated twitter and then 17 percent
mentioned instagram. Overall social media platforms Facebook, Instagram
and Twtitter were less popular methods of communciation. This could be
because the survey only included individals over the age of 18.
Moreover, the median sample was in the 40s.

@@

\caption{\label{fig-four}The stacked bar plot illustrates the preferred
methods of communication for receiving information about climate change
and climate action.}

@@

(\textbf{table-four?}) compares responses from 2018 and 2021 surverys
regarding the best method of delivering information on climate change
and climate action. Aside from the additon of four options to the 2021
survey, both surveys questions are the same. The leading category for
2018 is advertising campaigns with 55 percent, closely followed by
newsletter emails with 52 percent and the Toronto website with 46. The
2021 survey illustrates a consistent trend with the same platforms as
the top method of delivering information, specifically the Toronto
website with 44 percent, then enewsletter emails with 43 percent and
advertising campains at 42 percent. These top methods of communciation
are more evenly spread out than the 2018 survey results. Social media
platforms namely twitter, facebook and instagram all increased in the
2021 survey. Notable decreases in term of percentage from 2018 to 2021
include events, enewsletter emails, advertising campaigns, brochures and
pamphlets. The new category of BetterHomesTO.ca webiste accounts for 15
percent of votes for the 2021 survey.

@@

The likelihood of taking action against climate change was explored by
assessing individuals' intentions and behaviors. Figure~\ref{fig-five}
illustrates self-reported likelihood of taking certain actions that
minimize climate change effects. On the x-axis, we have the nine
different climate actions and on the y-axis is the percentage. The
stacked bar chart allows understanding of how likely individuals are to
take part in the given actions. Sorting waste into the correct bins is
the action that most individuals are doing right now, accounting for 60
percent whith 24 percent saying they are very likely to take part in
this action. The next most prominent actions are walking or cycling
short distances, reduce hydro usage, reduce waste and home improvements.
40 percent of invidiuals report walking or cycling when traveling
shorter distances in the city with 22 percent reporting they are very
likely. However, 7 percent say they are very unlikely to change their
method of transportation. Reducing the amount of energy and water use is
already done by 36 percent and very likely for 29 percent only a mere 3
percent is very unlikely. Reducing personal waste such as contributing
to repurposing used products has similar percentages to responses for
reducing hydro but even fewer individuals that report they are very
unlikely to take action. Similarly, the likelihoods for making home
improvements for energy efficiency such as energy-efficient appliances,
programmable thermostat, LED lightbulbs, green or cool roof. In
contrast, the action where the least number of individuals are
contributing to is the investment of an electric or hybrid vehicle over
gas-powered vehicles with only 8 percent who currently have an energy
efficient vehicle. While 19 percetn are very likely to make this switch,
14 percent say they are very unlikely. This illustrates that buying an
electric or hybrid vehicle is action the least number of individuals are
willing to invest in despite gas-powered vehicles being a major
contributor to greenhouse gases. This action has a huge impact on the
environment and yet 41 percent of this sample say they are unlikely to
change! The second action that individuals are most unlikely to engage
in is eat less meat by incorporating more plant-based foods in their
diet with 12 percent very unlikely and 23 percent somewhat unlikely.
Only 23 percent are already consuming meat alternatives. Vegans
greenhosue gas emissions are on average 37\% lower than those who eat
meat frequently.

@@

\caption{\label{fig-five}The stacked bar plot shows the likelihood of
survey respondents taking various actions to address climate change.}

@@

\subsubsection{Likelihood of investing in an electric or hybrid
vehicle}\label{likelihood-of-investing-in-an-electric-or-hybrid-vehicle}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/vehicle_electric_tree.png}

@@

\caption{Probability tree for likelihood of investing in an electric or
hybrid vehicle, predicted by age and education.}

@@

\subsubsection{Likelihood of making home
improvements}\label{likelihood-of-making-home-improvements}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/home_improvement_tree.png}

@@

\caption{Probability tree for likelihood of making home improvements
predicted by age and education.}

@@

\subsubsection{Likelihood of adopting meat
alternatives}\label{likelihood-of-adopting-meat-alternatives}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/protein_alternative_tree.png}

@@

\caption{Probability tree for likelihood of adopting meat alternatives,
predicted by age and education.}

@@

\subsubsection{Likelihood of minimizing car
use}\label{likelihood-of-minimizing-car-use}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/minimize_car_tree.png}

@@

\caption{Probability tree for likelihood of minimizing car use,
predicted by age and education.}

@@

\subsubsection{Likelihood of purchasing green
products}\label{likelihood-of-purchasing-green-products}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/green_product_tree.png}

@@

\caption{Probability tree for likelihood of purchasing green products,
predicted by age and education.}

@@

\subsubsection{Likelihood of reducing hydro
usage}\label{likelihood-of-reducing-hydro-usage}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/reduce_hydro_tree.png}

@@

\caption{Probability tree for likelihood of reducing hydro usage,
predicted by age and education.}

@@

\subsubsection{Likelihood of reducing
waste}\label{likelihood-of-reducing-waste}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/reduce_waste_tree.png}

@@

\caption{Probability tree for likelihood of reducing waste, predicted by
age and education.}

@@

\subsubsection{Likelihood of sorting waste
correctly}\label{likelihood-of-sorting-waste-correctly}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/sort_waste_tree.png}

@@

\caption{Probability tree for likelihood of sorting waste correctly,
predicted by age and education.}

@@

\subsubsection{Likelihood of walking or cycling shorter
distances}\label{likelihood-of-walking-or-cycling-shorter-distances}

\begin{figure}[H]

{\centering \includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/short_distance_tree.png}

@@

\caption{Probability tree for likelihood of walking or cycling shorter
distances, predicted by age and education.}

@@

The table presents the accuracy of models predicting the likelihood of
various climate-friendly actions based on demographic factors such as
age and education. Each row represents a different climate action (e.g.,
home improvement, reducing hydro, minimizing car use), with the
corresponding accuracy of the model's predictions. The accuracy values
reflect how well the models, which use age and education as predictors,
align with the actual data. An accuracy of 33 percent for home
improvement'' indicates that the model correctly predicted the outcome
33\% of the time for this action. These accuracy values help evaluate
the effectiveness of age and education as predictors for each specific
climate action.

@@

Figure~\ref{fig-six} shows the relationship between age and the
likelihood of engaging in various climate-related actions. The x-axis
represents the age groups, and the y-axis indicates the percentage of
respondents reporting different likelihood levels, from ``Already doing
this'' to ``Very unlikely.'' The trend lines across each action show how
engagement varies by age.

Younger respondents (15-24, 25-34) are less likely to adopt electric or
hybrid vehicles, with only a small proportion already using or
considering them. Adoption rates increase slightly in the 55-64 age
group, then decline again in the 65+ group. This suggests that older
individuals may have more resources or stability to invest in electric
vehicles.

Middle-aged individuals (35-54) are the most likely to engage in home
improvement actions aimed at increasing energy efficiency. Younger age
groups (15-24, 25-34) report lower likelihoods, likely due to barriers
such as homeownership and disposable income.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Predicting the Likelihood of Climate-Friendly Actions: Insights from Age and Educational Attainment},
  pdfauthor={Lexi Knight},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Predicting the Likelihood of Climate-Friendly Actions: Insights
from Age and Educational Attainment\thanks{Code and data are available
at: \url{https://github.com/LexiKnight/toronto_climate/tree/main}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{An analysis of how demographic factors influence the adoption
of sustainable behaviors}
\author{Lexi Knight}
\date{December 10, 2024}

@@

The \texttt{arrow} package was employed to handle large datasets
efficiently, ensuring smooth data processing, while \texttt{forcats} was
used for reordering factor levels to facilitate more insightful analyses
and visualizations. In terms of data tidying, the \texttt{tidyr} and
\texttt{stringr} packages allowed us to reshape and clean the data by
managing character strings and handling missing values. The (Therneau
and Atkinson 2024) and (Hothorn and Zeileis 2024) were used in to make
the models. For the final documentation and report generation, the
\texttt{knitr} package was instrumental in integrating R code and
outputs into a cohesive and reproducible report format. This suite of R
packages collectively enabled us to conduct thorough data management,
manipulation, and analysis, ultimately contributing to the successful
completion of the project.

@@

Figure~\ref{fig-two} illustrates the different age ditribution for 2018
and 2021 surveys. The major difference is that there was greater
representation of younger inviduals ranging from 18 to 23 in 2021
compared to 2018. The remaining age groups were not very different.

@@

\caption{\label{fig-two}This table highlights the age distribution of
survey respondents in 2018 and 2021.}

@@

Figure~\ref{fig-three} illustrates the distribution of survey
respondents by their highest level of education attained. The most
frequently reported education levels are ``Completed undergraduate
degree'' and ``Post graduate/professional school,'' with both categories
having a noticeably higher frequency compared to others, indicating a
well-educated respondent population. In contrast, fewer respondents
reported ``Some community college, vocational, trade school'' and
``Completed community college, vocational, trade school,'' suggesting
these levels are less common among participants. Additionally, a small
portion of respondents preferred not to disclose their education level.
This distribution highlights a skew toward higher educational attainment
within the sample, which may influence responses related to
socio-economic factors and other survey topics.

@@

\caption{\label{fig-three}The histogram shows the distribution of
education levels of 2018 survey respondents.}

@@

Figure~\ref{fig-four} illustrates respondant education level in 2018 and
2021 respectively. Please note, the categories of education level are
different for 2018 and 2021 as these were different surveys and
categories were updated in 2021. Despite this, we are still able to make
comparisons. Individuals with high school diplomas or less account for 9
for 2018 and 38 for 2021 (addition of less than highs school and high
school or equivalent categories). This could indicate that the younger
population was included more in the 2021 suvey or that less education
individuals were surveyed. Additionally, its not super clear howver
those with a post graduate or professional degree accounted for 29
percent in 2018 and 35 percent in 2021. This illustrates that more
educated inviduals responded to the 2021 survey.

@@

\caption{\label{fig-four}This table highlights the education
distribution of survey respondents in 2018 and 2021.''}

@@

Figure~\ref{fig-five} depicts the distribution of self-reported climate
change knowledge among survey respondents in 2018. The majority of
participants identified as ``Very informed,'' making this the most
prevalent category by a significant margin. A moderate number of
respondents reported being ``Not very informed,'' while a smaller yet
noticeable group described themselves as ``Extremely informed.'' In
contrast, the categories ``Not at all informed'' and ``Not very
informed'' have the lowest frequencies, with ``Not at all informed''
being particularly rare. This distribution suggests that most
respondents possess a relatively high level of awareness about climate
change, with very few indicating a lack of knowledge on the topic.

@@

\caption{\label{fig-five}The histogram shows the distribution of
self-reported climate change knowledge among 2018 survey respondents.}

@@

Figure~\ref{fig-six} illustrates the self-reported level of extent
inviduals feel informed about climate change and climate action for 2018
and 2021. In 2018, the large marjority of individuals say they are very
informed with 60 percent and the follwoign category is not very informed
with 27 percent. In 2021, The majority of inividuals also say they are
very informed but to a lesser degree with 48 percent and this is closely
followed by only a little informed with 37 percent. There is a trend
that invidiusls in 2018 reported greater levels of knowldege about
climate change whereas 2021 illstrates a more even split between very
informed and only little informed, portraying that fewer individuals
have adequate knowledge about climate change.

@@

\caption{\label{fig-six}This table highlights the extent feel informed
distribution of survey respondents in 2018 and 2021.}

@@

T@ fig-seven, shows the analysis of the best method for delivering
climate change information portrayed in Figure~\ref{fig-four} revealed
that the majority of individuals in the 2018 dataset preferred receiving
information firstly from advertising campaigns with about 55 percent,
newsletters with about 52 percent, followed by the toronto website wiht
about 46 percent. On the other side of the spectrum, few said they were
uninterested in receiving climate action information with about 5
percent, about 12 percent indicated twitter and then 17 percent
mentioned instagram. Overall social media platforms Facebook, Instagram
and Twtitter were less popular methods of communciation. This could be
because the survey only included individals over the age of 18.
Moreover, the median sample was in the 40s.

@@

\caption{\label{fig-seven}The stacked bar plot illustrates the preferred
methods of communication for receiving information about climate change
and climate action.}

@@

\textbf{?@fig-eight} compares responses from 2018 and 2021 surverys
regarding the best method of delivering information on climate change
and climate action. Aside from the additon of four options to the 2021
survey, both surveys questions are the same. The leading category for
2018 is advertising campaigns with 55 percent, closely followed by
newsletter emails with 52 percent and the Toronto website with 46. The
2021 survey illustrates a consistent trend with the same platforms as
the top method of delivering information, specifically the Toronto
website with 44 percent, then enewsletter emails with 43 percent and
advertising campains at 42 percent. These top methods of communciation
are more evenly spread out than the 2018 survey results. Social media
platforms namely twitter, facebook and instagram all increased in the
2021 survey. Notable decreases in term of percentage from 2018 to 2021
include events, enewsletter emails, advertising campaigns, brochures and
pamphlets. The new category of BetterHomesTO.ca webiste accounts for 15
percent of votes for the 2021 survey.

@@

The likelihood of taking action against climate change was explored by
assessing individuals' intentions and behaviors. Figure~\ref{fig-nine}
illustrates self-reported likelihood of taking certain actions that
minimize climate change effects. On the x-axis, we have the nine
different climate actions and on the y-axis is the percentage. The
stacked bar chart allows understanding of how likely individuals are to
take part in the given actions. Sorting waste into the correct bins is
the action that most individuals are doing right now, accounting for 60
percent whith 24 percent saying they are very likely to take part in
this action. The next most prominent actions are walking or cycling
short distances, reduce hydro usage, reduce waste and home improvements.
40 percent of invidiuals report walking or cycling when traveling
shorter distances in the city with 22 percent reporting they are very
likely. However, 7 percent say they are very unlikely to change their
method of transportation. Reducing the amount of energy and water use is
already done by 36 percent and very likely for 29 percent only a mere 3
percent is very unlikely. Reducing personal waste such as contributing
to repurposing used products has similar percentages to responses for
reducing hydro but even fewer individuals that report they are very
unlikely to take action. Similarly, the likelihoods for making home
improvements for energy efficiency such as energy-efficient appliances,
programmable thermostat, LED lightbulbs, green or cool roof. In
contrast, the action where the least number of individuals are
contributing to is the investment of an electric or hybrid vehicle over
gas-powered vehicles with only 8 percent who currently have an energy
efficient vehicle. While 19 percetn are very likely to make this switch,
14 percent say they are very unlikely. This illustrates that buying an
electric or hybrid vehicle is action the least number of individuals are
willing to invest in despite gas-powered vehicles being a major
contributor to greenhouse gases. This action has a huge impact on the
environment and yet 41 percent of this sample say they are unlikely to
change! The second action that individuals are most unlikely to engage
in is eat less meat by incorporating more plant-based foods in their
diet with 12 percent very unlikely and 23 percent somewhat unlikely.
Only 23 percent are already consuming meat alternatives. Vegans
greenhosue gas emissions are on average 37\% lower than those who eat
meat frequently.

@@

\caption{\label{fig-nine}The stacked bar plot shows the likelihood of
survey respondents taking various actions to address climate change.}

@@

\subsubsection{Decision Trees}\label{decision-trees}

A decision tree is a machine learning model used to make predictions by
recursively splitting data into subsets based on specific conditions of
the predictor variables. It is structured like a tree, where the top
node is the root node, representing the entire dataset, and subsequent
nodes represent splits based on decision rules. The branches indicate
the outcomes of these rules, and the leaf nodes (terminal nodes) provide
the final prediction for each subset.

Each split is determined by the predictor variable most strongly
associated with the target outcome, maximizing predictive accuracy. A
split separates individuals by age or educational level. The prediction
in each terminal node includes the majority class (e.g., ``Very
likely'') and additional information such as sample size (n) and error
percentage. The error percentage reflects the proportion of cases in
that node misclassified by the model. A higher error percentage
indicates greater uncertainty or variability within the group, while a
lower error percentage suggests a more precise classification.

\paragraph{Likelihood of investing in an electric or hybrid
vehicle}\label{likelihood-of-investing-in-an-electric-or-hybrid-vehicle}

Figure~\ref{fig-ten} predicts individuals' likelihood to invest in an
electric or hybrid vehicle based on their highest level of education and
age. The model identifies highest level of education as the most
important variable for the initial split. At the root node, individuals
are divided into two groups: those with ``Some Community/Trade School''
or ``Some University'' education and those with other educational
backgrounds such as ``High School,'' ``Undergrad,'' ``Postgrad,'' or
``Prefer Not to Answer.''

For individuals with ``Some Community/Trade School'' or ``Some
University'' education, the tree predicts that they are ``Very likely''
to invest, with a sample size of 47 and an error rate of 63.8\%. This
indicates moderate variability within this group.

For those with other educational backgrounds, further splits are based
on age. Younger individuals (below 38.5 years) are segmented further. A
subset aged under 34.5 years is split into two groups, with one group (n
= 10) being ``Very likely'' to invest, with a lower error rate of 30\%.
The other group is classified as ``Somewhat likely,'' with a higher
error rate of 60.7\%. Older individuals (38.5 years or older) are less
likely to invest. For instance, those aged above 68.5 years are
predicted as ``Very unlikely'' with a sample size of 18 and an error
rate of 61.1\%, while those aged between 34.5 and 68.5 years are divided
into ``Somewhat unlikely'' or ``Somewhat likely'' categories.

Overall, the decision tree reveals distinct patterns where educational
attainment and age interact to predict investment likelihood, with
younger individuals and those with less formal education being more
likely to invest.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/vehicle_electric_tree.png}

@@

\caption{\label{fig-ten}Probability tree for likelihood of investing in
an electric or hybrid vehicle, predicted by age and education.}

@@

\paragraph{Likelihood of making home
improvements}\label{likelihood-of-making-home-improvements}

In Figure~\ref{fig-eleven}, the root node splits the data at age 35.5,
indicating that age is the most influential factor in determining
climate-related home improvement behavior. Individuals aged 36 years or
older are predominantly classified as ``Already doing this or have done
this,'' with a large sample size (n = 223) but a moderate error rate of
59.6\%, reflecting variability in this group.

For individuals younger than 36 years, the tree introduces education as
the next important factor. Those with higher educational attainment
(``Community/Trade School,'' ``Some University,'' or ``Postgraduate'')
are classified as ``Very likely'' to make climate-conscious home
improvements (n = 40) with an error rate of 55.0\%. Conversely,
individuals with lower education levels, such as ``High School'' or
``Some Community/Trade School,'' are classified as ``Somewhat likely''
(n = 46) but have a higher error rate of 67.4\%, indicating less
predictive certainty. A smaller subset of younger individuals with
similar educational attainment is classified as ``Somewhat likely'' (n =
10) with a lower error rate of 40.0\%, suggesting higher confidence in
this classification. The decision tree highlights clear trends based on
age and education, with older age groups and higher educational
attainment linked to greater engagement in making home improvements.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/home_improvement_tree.png}

@@

\caption{\label{fig-eleven}Probability tree for likelihood of making
home improvements predicted by age and education.}

@@

\paragraph{Likelihood of adopting meat
alternatives}\label{likelihood-of-adopting-meat-alternatives}

Figure~\ref{fig-twelve} predicts individuals' likelihood of consuming
meat alternatives to minimize the effects of climate change based on
their age and highest level of education. The model identifies age as
the most important variable for the initial split. At the root node,
individuals are divided into two groups: those under 34.5 years old and
those 34.5 years or older. This highlights age as a critical factor
influencing the likelihood of adopting meat alternatives.

In the subgroup of individuals under 34.5 years old, the decision tree
further splits based on age at 31.5 years, classifying younger
individuals into categories of ``Somewhat likely'' (n = 25, error =
44\%) and ``Somewhat unlikely'' (n = 62, error = 62.9\%). On the other
hand, for individuals aged 34.5 years or older, the model considers
highest level of education as the next key variable, separating those
with community/trade school, undergraduate, and postgraduate education
from those with high school education. This distinction reflects how
educational attainment moderates the impact of age on the decision to
adopt meat alternatives.

Within the older age group, additional splits occur based on age, such
as at 49 years and 65.5 years, leading to further classifications like
``Already doing this or have done this'' or ``Somewhat likely.'' These
splits indicate how the interaction between age and education level
shapes patterns of meat alternative adoption, resulting in more granular
predictions.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/protein_alternative_tree.png}

@@

\caption{\label{fig-twelve}Probability tree for likelihood of adopting
meat alternatives, predicted by age and education.}

@@

\paragraph{Likelihood of minimizing car
use}\label{likelihood-of-minimizing-car-use}

Figure~\ref{fig-thirteen} predicts individuals' likelihood of minimizing
car use as a measure to mitigate climate change. In this case, the tree
structure is straightforward, containing only a single root node. The
model classifies all individuals into one category, ``Already doing this
or have done this,'' with a sample size of 320 and an error rate of
61.9\%. The absence of splits in the tree indicates that no other
variables, such as age or education, contributed sufficient predictive
power to justify further segmentation.

This outcome suggests that either the input data lacks variability in
relevant features, or the behavior of minimizing car use is widespread
and consistent across the sample population. Consequently, the model
does not differentiate between subgroups and treats all individuals as
having the same likelihood of engaging in this behavior.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/minimize_car_tree.png}

@@

\caption{\label{fig-thirteen}Probability tree for likelihood of
minimizing car use, predicted by age and education.}

@@

\paragraph{Likelihood of purchasing green
products}\label{likelihood-of-purchasing-green-products}

\textbf{?@fig-fourteen} predicts individuals' likelihood of purchasing
green products to minimize the effects of climate change based on age
and highest level of education. The model identifies age as the most
important variable for the initial split. At the root node, individuals
are divided into two groups: those younger than 37.5 years old and those
37.5 years or older. This division highlights age as a key determinant
in predicting green product purchasing likelihood.

For individuals aged 37.5 years or older, the next split is determined
by the highest level of education. Those with educational backgrounds
such as ``High School,'' ``Postgrad,'' ``Some Community/Trade School,''
``Some University,'' or ``Prefer Not to Answer'' are further analyzed.
Within this group, additional splits are made based on age thresholds of
50.5 years and 47.5 years. These splits refine the classification into
categories such as ``Somewhat likely'' or ``Already doing this or have
done this,'' with varying error rates and sample sizes.

Among individuals younger than 37.5 years, the highest level of
education again plays a crucial role in the split. Individuals with
educational backgrounds such as ``Postgrad,'' ``Community/Trade
School,'' ``High School,'' or ``Some University'' are further segmented.
Within this subgroup, age thresholds of 27.5 years and 22.5 years are
used to differentiate individuals into categories such as ``Very
likely'' or ``Somewhat likely,'' reflecting diverse levels of engagement
with green product purchasing. Overall, the decision tree incorporates
age and education as the primary variables to segment individuals into
likelihood categories.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/green_product_tree.png}

@@

\caption{\label{fig-forteen}Probability tree for likelihood of
purchasing green products, predicted by age and education.}

@@

\paragraph{Likelihood of reducing hydro
usage}\label{likelihood-of-reducing-hydro-usage}

Figure~\ref{fig-fifteen} identifies age and highest level of education
as key predictors influencing the likelihood of reducing hydro usage.
The root node of the tree reveals that individuals younger than 34.5
years old are categorized as ``Very likely'' to reduce hydro usage, with
a sample size of 87 and an error rate of 60.9\%. This suggests that
younger individuals may demonstrate a greater propensity to adopt
hydro-reduction practices. Conversely, individuals over 34.5 years are
further split by age, with those older than 64.5 years also categorized
as ``Very likely,'' albeit with a smaller sample size of 16 and a lower
error rate of 37.5\%. This bifurcation highlights that age groups at the
extremes of the spectrum (younger and older individuals) are more
inclined toward hydro-use reduction.

Among middle-aged individuals (34.5--64.5 years), the decision tree
highlights the role of education. Those younger than 61.5 years and
possessing higher levels of education, including university degrees, are
either ``Already doing this or have done this'' (n=151, err=56.3\%) or
``Somewhat likely'' (n=14, err=50.0\%) to engage in hydro-use reduction.
On the other hand, individuals with lower educational attainment,
particularly trade school or community-level education, are less likely
to implement such measures. For individuals aged 61.5--71.5 years, there
remains a balance between those already reducing hydro usage and those
who are very likely to adopt it in the future, with clear variability
based on educational level.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/reduce_hydro_tree.png}

@@

\caption{\label{fig-fifteen}Probability tree for likelihood of reducing
hydro usage, predicted by age and education.}

@@

\paragraph{Likelihood of reducing
waste}\label{likelihood-of-reducing-waste}

Figure~\ref{fig-sixteen} predicts the likelihood of reducing waste is
based on age and highest level of education. The root node splits
individuals by age at 34.5 years, indicating that younger and older
individuals exhibit distinct waste-reduction behaviors. For those under
34.5 years, the model uses highest level of education as a key
determinant, classifying them as either ``Somewhat likely'' or ``Very
likely'' to reduce waste.

For individuals aged 34.5 years or older, the tree splits again by
education and further refines predictions using additional age
thresholds at 61.5 and 75.5 years. Older individuals with higher
education levels are more likely to have ``Already done this or are
doing this,'' while those with lower education levels are categorized as
``Somewhat likely'' or ``Very likely.'' This structure highlights the
combined influence of education and age on waste-reduction behaviors,
with clear thresholds that differentiate groups.

The decision tree demonstrates that age is the primary driver of
waste-reduction behavior, with education acting as an important
secondary factor. These variables together capture differences in
sustainability habits, showing how younger, middle-aged, and older
populations vary in their likelihood to engage in reducing waste.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/reduce_waste_tree.png}

@@

\caption{\label{fig-sixteen}Probability tree for likelihood of reducing
waste, predicted by age and education.}

@@

\paragraph{Likelihood of sorting waste
correctly}\label{likelihood-of-sorting-waste-correctly}

Figure~\ref{fig-seventeen} predicts the likelihood of sorting waste
correctly identifies highest level of education as the primary
determinant. Individuals with education levels at or above
Community/Trade School, Postgraduate, or Some University are categorized
as ``Already doing this or have done this,'' accounting for the majority
of respondents (n = 295, err = 37.3\%). This demonstrates that higher
educational attainment correlates with proactive waste-sorting
behaviors.

For individuals whose highest level of education is High School, the
model introduces a secondary split based on age at 37 years. Younger
individuals under 37 years are categorized as ``Already doing this or
have done this,'' albeit with a higher error rate (n = 7, err = 57.1\%).
Older individuals aged 37 or above are more likely classified as ``Very
likely'' to sort waste correctly (n = 17, err = 35.3\%). These splits
highlight the combined influence of education and age on waste-sorting
behavior, with specific subgroups showing varying levels of commitment
and accuracy.

This decision tree underscores that education plays a dominant role in
predicting waste-sorting behavior, while age further refines these
predictions, especially for those with lower education levels. The model
provides straightforward insights into how demographics shape
sustainable practices.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/sort_waste_tree.png}

@@

\caption{\label{fig-seventeen}Probability tree for likelihood of sorting
waste correctly, predicted by age and education.}

@@

\paragraph{Likelihood of walking or cycling shorter
distances}\label{likelihood-of-walking-or-cycling-shorter-distances}

Figure~\ref{fig-eighteen} predicts the likelihood of individuals walking
or cycling short distances based on their highest level of education and
age. The root node identifies education level as the most significant
factor. Individuals with postgraduate, undergraduate, or some university
education are classified as ``Already doing this or have done this'' (n
= 231, error = 51.9\%). This indicates a strong relationship between
higher education levels and the adoption of walking or cycling short
distances.

For those with lower education levels (community/trade school, high
school, or preferring not to answer), the model further splits based on
age. The first division occurs at 65.5 years, distinguishing those
``Very likely'' to walk or cycle short distances (n = 15, error = 60\%)
from others. Additional splits by age at 33.5 years and 60.5 years,
along with education, refine the classification. This stepwise approach
demonstrates how age interacts with education to influence
short-distance mobility choices.

\includegraphics[width=10.67in,height=\textheight]{../data/04-model_data/short_distance_tree.png}

@@

\caption{\label{fig-eighteen}Probability tree for likelihood of walking
or cycling shorter distances, predicted by age and education.}

@@

\textbf{?@fig-nineteen} presents the accuracy of models predicting the
likelihood of various climate-friendly actions based on demographic
factors such as age and education. Each row represents a different
climate action (e.g., home improvement, reducing hydro, minimizing car
use), with the corresponding accuracy of the model's predictions. The
accuracy values reflect how well the models, which use age and education
as predictors, align with the actual data. An accuracy of 33 percent for
home improvement'' indicates that the model correctly predicted the
outcome 33\% of the time for this action. These accuracy values help
evaluate the effectiveness of age and education as predictors for each
specific climate action.

@@

Figure~\ref{fig-twenty} shows the relationship between age and the
likelihood of engaging in various climate-related actions. The x-axis
represents the age groups, and the y-axis indicates the percentage of
respondents reporting different likelihood levels, from ``Already doing
this'' to ``Very unlikely.'' The trend lines across each action show how
engagement varies by age.
